%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Document: Thesis for PhD at UC Riverside%%
%% Title: Investigating the evolution of environmental and biotic interactions %%
%%  in basal fungal lineages through comparative genomics  %%
%% Autnor: Steven Ahrendt  %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% INHIBITION CHAPTER %%
%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Growth suppression properties of the Chytridiomycete \textit{Homolaphlyctis polyrhiza} JEL142}
\label{chap:HpInhibition}
%%%%%%%%%%%%%%%%%%
%% INTRODUCTION %%
%%%%%%%%%%%%%%%%%%
\section{Introduction}
Interactions between microorganisms are facilitated by biological signals. These include proteins, small molecules, and various chemical compounds, either bound to the cell surface or secreted into the environment. Many of these compounds can be classified as secondary metabolites: chemicals not required for growth or development of the organism. A number of well-known naturally-derived antifungal products, including toxins \cite{Kokkonen2010}, antimicrobials \cite{Wiemann2014}, and plant hormone mimics \cite{Howlett2006}, can be classified as secondary metabolites. Secretion by an organism in a resource-limited environment of secondary metabolites which also happen to negatively impact neighboring organisms would confer a selective advantage upon the producer. Thus, resource competition likely plays a role in the evolution of natural antifungal production \cite{Vicente2003}.\\
\indent In humans, the most common fungal pathogen is \textit{Candida albicans} (Ascomycota) \cite{Sherris1994}. It is a commensal of healthy gut flora, but overgrowth results in opportunisit infection and the disease candidiasis. \textit{Cryptococcus neoformans} (Basidiomycota) is another common human pathogen. It is typically found in soil and associated with pigeon guano but causes severe meningitis in patients with AIDS or others with similarly compromised immune systems \cite{Casadevall2013}.\\
\indent After surveying the international fungal pathology community regarding plant pathogens, the filamentous Ascomycete \textit{Magnaporthe oryzae} was deemed to be the most important in terms of scientific/economic merit \cite{Dean2012}. \textit{M. oryzae} is the causative agent of rice blast disease, a devastating disease given that roughly one-half of the global population is reliant on rice. As such, this fungus has developed into model system for studying plant-pathogen interactions. The Ascomycete \textit{Botrytis cinerea}, also known as grey mould, was deemed to be the second most important fungal plant pathogen. Due to its broad host range which includes over 200 species, it is the most extensively studied necrotrophic fungal pathogen \cite{Dean2012}.\\
\indent Other important fungal plant pathogens infect wheat (\textit{P. graminis} f. sp. \textit{tritici}, causes black rust; \textit{P. striiformis} f. sp. \textit{tritici}, causes yellow rust; \textit{P. triticina}, causes brown rust), cereals (\textit{Fusarium graminearum}, \textit{Blumeria graminis}, \textit{Mycosphaerella graminicola}), tomatoes (\textit{Fusarium oxysporum}), and corn (\textit{Ustilago maydis}).\\
\indent Therefore, research into new natural antifungal treatment is increasingly important as these and other pathogenic fungi become progressively resistant to current, conventionally synthesized fungicides and antifungal drugs (see: \cite{Bossche1998}, \cite{Kontoyiannis2002}, \cite{Kretschmer2009}, \cite{Leroch2011}, and \cite{Pfaller2012}). The bulk of research into fungal-derived secondary metabolites has thus far focused primarily on Ascomycete and Basidiomycete fungi due to their diverse product types \cite{Berdy2012}. Chytrids are a relatively understudied group for secondary metabolites and represent unexplored diversity which may harbor novel compounds and molecules. Comparative genomic analyses have identified a variety of degradation enzymes, but a detailed examination of secondary metabolites involved in competition zoosporic fungi has yet to be explored. It is likely that these organisms have evolved mechanisms for mediating chemical interactions with other organisms, and therefore could present a potentially valuable source of novel antifungal compounds.\\
\indent \textit{Homolaphlyctis polyrhiza} (\textit{Hp}) is a non-pathogenic member of the Chytridiomycota and is most closely related to the amphibian pathogen \textit{B. dendrobatidis} (\textit{Bd}). The specific isolate, \textit{Hp} JEL 142, has been used previously in phylogenetic studies of chytrids \cite{James2000,James2006sixGene,Letcher2008} and was provided a formal name in 2011 \cite{Longcore2011}. A draft 454 genome assembly was produced in that same year in order to gain comparative insights into pathogenicity of \textit{Bd} \cite{Joneson2011}.\\
\indent This isolate of \textit{Hp} was collected by Dr. Joyce Longcore's research group in Maine, USA from a 1.7 ha oligotrophic and fishless lake with a pH of 4.6 \cite{Davis1994,Rhodes1995}. It was cultured using onionskin bait and isolated into pure culture on mPmTG nutrient agar. Its overall morphology is typical of members of the Rhizophydiales \cite{Barr1978} and includes a single lobed mitochondrion, a single libpid globule, microbody and rumposome placement, and Kinetosome / nonflagellated centriole ultrastruture. Phylogenetic reanalysis based on maximum likelihood using rRNA subunits established the placement of \textit{Hp} as sister to \textit{Bd}, and prompted the description of a novel genus (\textit{Homolaphlyctis}, in recognition of the researchers who helped with collection and analysis) and species (\textit{H. polyrhiza}, in reference to its multiple rhizoidal axes) \cite{Longcore2011}.\\
\indent While working with \textit{Hp} in the Stajich lab for the phototaxis and protein expression experiments described in Chapter~\ref{chap:RhodAux}, unintentional contamination of \textit{Hp} plates by \textit{Neurospora crassa} led to the unique observation of growth inhibition by a chytrid on the filamentous ascomycete. Specifically, I observed a distinctive zone created by \textit{N. crassa} hyphae growing near \textit{Hp} sporangia, while \textit{N. crassa} contamination on plates of related Chytridiomycetes \textit{Batrachochytrium dendrobatidis} (\textit{Bd}) and \textit{Spizellomyces punctatus} (\textit{Sp}) did not result in the same zone of inhibition.\\
\indent This property of \textit{Hp} is reliable and reproducible. This chapter describes the experimental and computational work towards identification and characterization of a compound responsible for hyphal growth suppression of \textit{N. crassa} by \textit{Hp}. This study is the first of \textit{Hp} in the context and presents novel findings about microbial interactions within early branching fungal lineages. Assistance in preparing media and general strain maintenance was provided by undergraduate students in the Stajich lab, Na Jeong and Sapphire Ear, and visiting student in the Research Experience for Undergraduates (REU) program Spencer Swansen.\\

%%%%%%%%%%%%%
%% Methods %%
%%%%%%%%%%%%%
\section{Methods}

%---- Growth of cultures ----%
\subsection*{Fungal strains and maintenance}
Cultures of \textit{Homolaphlyctis polyrhiza} JEL142, \textit{Operculomyces laminatus} JEL223, \textit{Rhizoclosmatium hyalinus} JEL800, and \textit{Obelidium mucronatum} JEL802 were individually maintained on mPmTG [peptonized milk (0.4 g/L), tryptone (0.4 g/L), dextrose (2 g/L)], \textit{Batrachochytrium dendrobatidis} JEL423 cultures were maintained on 1\% Tryptone [tryptone (10 g/L), dextrose (3.2 g/L)], and \textit{Spizellomyces punctatus} SW-1 cultures were maintained on PmTG [peptonized milk (0.5 g/L), tryptone (1 g/L), dextrose (5 g/L)]. All chytrid cultures were maintained at room temperature (approx. 23$^{\circ}$C). Unless otherwise specified, all experiments using chytrids were carried out using the specific media described above for each species. Motile chytrid zoospores, when required, were obtained from actively growing (ie 2-4 day old) plates by flooding and subsequently (after 30-45 minutes at room temperature) collecting 2-4 mls of sterile di H$_{2}$O. Sporangia samples for inoculation were obtained by removing a block (approx. 1 cm$^{2}$) of agar containing actively-growing chytrid sporangia.\\
\indent Vogel's minimal medium (VM) \cite{Davis1970} was used for vegetative growth of \textit{Neurospora crassa} FGSC 2489, \textit{Neurospora discreta} FGSC 8579 and \textit{Neurospora tetrasperma} FGSC 2508. \textit{N. crassa} kinase/phosphotase mutants (Table~\ref{tab:ChInhib_Kinase}; \cite{Park2011}) were maintained on VM + Hygromycin and grown similarly to \textit{N. crassa} WT strains. \textit{Trichoderma reesei} FGSC 10290, \textit{Phycomyces blakesleeanus}, and \textit{Ashbya gossypii} cultures were maintained on PDA media [potato dextrose agar (39 g/L)] at 28$^{\circ}$C, 20$^{\circ}$C, and 30$^{\circ}$C, respectively. \textit{Aspergillus nidulans} FGSC A4 cultures were maintained on minimal media [dextrose (10 g/L), nitrate salts (50 mL/L), Trace elements (1 ml/L)], at 28$^{\circ}$C. \textit{Saccharomyces cerevisiae} strains MAU99 and AH109 were maintained on YPDA [yeast extract (10 g/L), peptone (20 g/L), dextrose (20 g/L), adenine sulfate (0.003\%)] media at 35$^{\circ}$C. \textit{Coprinopsis cinerea} FGSC 9003 cultures were maintained on YPD media at 37$^{\circ}$C.  Plates and slants of 1\% tryptone, PmTG, mPmTG, VM, YPD, and PDA media included 1\% agar. Plates and slants of MM included 1.8\% agar. Plates and slants of YPDA included 2\% agar.\\

%---- Solid and liquid test ----%
\subsection*{Bioactivity on solid media and liquid media}
To assess the breadth of fungal species susceptible to \textit{Hp} sporangia, \textit{Hp} was individually co-cultured on mPmTG media with the selection of species described above. A block of agar containing active \textit{Hp} sporangia was added to mPmTG plate, flushed with sterile diH$_{2}$O, and incubated for 48h, at which point they were inoculated at a single point with 1x10$^{6}$ \textit{Neurospora} conidia in diH$_{2}$O. For \textit{Neurospora} experiments, plates were left to grow for an additional 24 - 48 hrs.\\
\indent Similarly, for other fungi (which are much slower growing than \textit{Neurospora}), \textit{Hp} and the target fungus were inoculated at the same time on an mPmTG solid media plate. Growth was monitored every other day until the other fungi reached \textit{Hp} sporangia, approximately 8-10 days.\\
\indent To obtain aliquots of filter-sterilized\textit{Hp}-conditioned media ("filtrate"), initial preparations (in triplicate) were constructed by adding eight blocks of agar containing actively-growing \textit{Hp} sporangia, each approximately 1cm$^{2}$ to 10 ml of liquid mPmTG. Initially, to determine if the compound was being constitutively produced or is a response mechanism to the presence of another organism, a second preparation was made identically to the one just described, to which an additional 0.5 $\mu$l of a 100 $\mu$l  \textit{N. crassa} conidial suspension was added. To control for possibility that growth rate changes could be caused by nutrient depletion by \textit{N. crassa} in this second trial, a third preparation was made with eight mPmTG agar blocks with no \textit{Hp} sporangial growth in 10 ml mPmTG, to which 0.5 $\mu$l of a 100 $\mu$l \textit{N. crassa} conidial suspension was added. These cultures preparations were left to incubate at room temperature for 72hrs, at which point the total filtrate was obtained by passing each individual replicate through a 0.22 $\mu$m syringe filter into a 50 ml conical tube. All replicates for a given treatment were pooled after filtration. \\
\indent Large scale preparations were constructed in a similar manner by adding twenty blocks of agar containing \textit{Hp} sporangia to 50 ml of mPmTG media, allowed to grow for 72hrs and filtered using a 0.22 $\mu$m syringe filter. A dilution series was prepared using twenty-five blocks of agar in 75 ml of mPmTG media and filtered using a 0.22 $\mu$m Stericup filter (Millipore). Conditioned \textit{Hp} media filtrate was used to dilute VM media at values of 0\%, 10\%, 25\%, 50\%, 75\%, 90\%, and 100\% filtrate, achieving a total volume of 3 ml in which 1x10$^{5}$ \textit{N. crassa} conidia were inoculated. Similarly, additional dilutions were prepared using fresh mPmTG media, \textit{Bd} filtrate (prepared similar to \textit{Hp} described above), and fresh 1\% Tryptone media.\\
\indent For temperature assays, the filtrate from each small scale preparation was subsequently divided into eight 2 ml aliquots (in duplicate) to test for bioactivity over three separate experiments. The control sample was left undisturbed and uninoculated throughout the experiment. Another sample was inoculated with 2x10$^{6}$ conidia but otherwise untreated. The filtrate was subjected to six different temperature treatments to test its thermostability: -80$^{\circ}$C (30 min), -20$^{\circ}$C (1 hr), 4$^{\circ}$C (1 hr), 28$^{\circ}$C (1 hr), 65$^{\circ}$C (1 hr), and 90$^{\circ}$C (30 min). Each experimental treatment was inoculated with 2x10$^{6}$ conidia. These temperature treatments were allowed to return to room temperature prior to inoculation.\\
\indent To test the sensitivity of \textit{Bd} to \textit{Hp}, 3 ml of the filtrate from the large scale preparation was used as an incubation medium for one blcok of 1\% Tryptone agar containing actively growing \textit{Bd} sporangia. Similarly, 3 ml of mPmTG media was used as a positive control. These preparations were incubated for 96h at room temperature to allow for potential zoospore production and release. After 96h, the suspension was mixed briefly and 1 ml was added to mPmTG plates, while another 1 ml was added to 1\% Tryptone plates. The plates were incubated at room temperature for a maximum of 14d and photographed periodically.\\

%---- Reassembly and annotation ----%
\subsection*{Reassembly and annotation}
\textit{Hp} sporangia material was collected by scraping actively growing \textit{Hp} plates with sterile spatulas. Genomic DNA was extracted using a modified bead-beating procedure. Briefly, approximately 100 mg of material containing both zoospores and sporangia was scraped from actively-growing, zoospore-rich \textit{B. dendrobatidis} and \textit{S. punctatus} plates. The material was added to approximately 100 mg of silicon beads (0.5mm dia.) and mixed with 600 $\mu$l of Cell lysis solution (Qiagen, Germantown, MD) and 3 $\mu$l of proteinase K. The solution was homegenized with a bead beater using a 30s pulse at 4$^{\circ}$C and subsequently incubated for 1.5h at 55$^{\circ}$C. 200 $\mu$l of protein precipitation solution (Qiagen, Germantown, MD) was added to the mixture and iced for 15 min. After centrifugation at 14000xg for 3 min at room temperature, the supernatant was collected, mixed with 600 $\mu$l isopropanol, and spun at 1400xg for 1 min at room temperature. Pellet was washed with 600 $\mu$l ice cold 70\% EtOH and spun at 1400xg for 1 min at room temperature. Finally the pellet was air dried for 15 min at room temperature, resuspended in 50 $\mu$l H$_{2}$O, incubated at 65$^{\circ}$C for 1 hr, and stored at -20$^{\circ}$C. A DNAseq Illumina library was prepared using the NEBNext Ultra DNA library kit and submitted to the University of California, Riverside Genomics Institute for Integrateve Genomic Biology (IIGB) core facility for MiSeq Illumina HT sequencing.\\
\indent Whole genome assembly was performed using the Celera Assembler \cite{Myers2000}. HT sequencing yielded 27,541,671 reads. The final assembly is approximately 27.3 Mb and consists of 2720 contigs (N50=22,735). The GC content is 48.2\%. For comparision, the previous 454 assembly utilized 922,085 reads producing 16,311 contigs (N50=36,162) and an assembly size of 26.7 Mb.\\
\indent Genome annotation was performed using MAKER (v2.31.8) \cite{Cantarel2008}, incorporating EST evidence from \textit{Batrachochytrium dendrobatidis}, and including RepeatMasker (v4.0.5), SNAP (v0.15.4), Augustus (v2.7), and Fgenesh. A total of 8155 proteins were predicted.\\

%---- in silico screen ----%
\subsection*{Secretome and small-metabolite prediction}
Putative secondary metabolite clusters were predicted from the annotated proteomes of \textit{Bd, Sp, Am, Ca}
and transcriptome from \textit{Cl} using antiSMASH \cite{Blin2013} and SMURF \cite{Khaldi2010}.\\
\indent To identify putative secretome proteins, these proteomes were also subjected to
predictive workflow optimized for fungi \cite{Min2010}. The workflow described therein
evaluated the prediction accuracies for a number of programs for eukaryotic secretome prediction:
SignalP \cite{Bendtsen2004}, Phobius \cite{Kall2004,Kall2007}, TargetP
\cite{Emanuelsson2000,Emanuelsson2007}, and WolfPSort \cite{Sprenger2006,Horton2007}.
These programs were evaluated indiviually and in combination with TMHMM \cite{Krogh2001} and PS-Scan
\cite{deCastro2006} and ranked by Mathews' Correlation Coefficient (MCC) \cite{Matthews1975,Baldi2000,Menne2000}
which is a measure of Sensitivity and Specificity. The fungal prediction workflow with
the highest accuracy (MCC: 84\%) combined SignalP/TMHMM/WolfPsort/Phobius/PS-Scan, in that order.
Here, I used SignalP (v4.1), WolfpSort (v0.2), and Phobius (v1.01) in the specific order given above to predict secreted proteins from
the proteomes for chytrids listed in Table~\ref{tab:AppData_taxa}. TMHMM (v2.0) and PS-Scan (v1.79) were
used to filter out transmembrane and ER targeting proteins, respectively.\\

%---- Temperature / pH growth ----%
\subsection*{Temperature and pH growth assays}
To compare to the assays described in \cite{Piotrowski2004}, \textit{Hp} was grown at a range of temperatures: 18$^{\circ}$C, 23$^{\circ}$C, 25$^{\circ}$C, 28$^{\circ}$C, and 30$^{\circ}$C. Briefly, a 15 ml mPmTG inoculum was incubated at 23$^{\circ}$C for 1 week. 1 ml of culture was used to inoculate 30 ml of mPmTG in 50 ml conical tubes. \\
\indent Similarly, \textit{Hp} and \textit{Bd} were grown at a range of pHs by preparing mPmTG and 1\% Tryptone media at pH levels: 4.0, 5.0, 6.0, 6.8, 8.0, and 9.0. Low-pH media were prepared with the addition of 1M HCl, while high-pH media were prepared with addition of 1M NaCl. pH levels were checked with a pH meter prior to autoclaving.\\

%%%%%%%%%%%%%
%% Results %%
%%%%%%%%%%%%%
\section{Results}

%---- Observation ----%
\subsection*{\textit{Hp} inhibition is unique among the chytrids and broadly active against other fungal species}
My initial observation was repeatable, and under controlled circumstances, I was able to demonstrate that \textit{Hp} sporangia are reliably and repeatedly capable of inhibition of \textit{N. crassa} vegetative growth on solid mPmTG media (Figure~\ref{fig:ChInhib_NcraChytrid}A). The other tested Chytridiomycota species, \textit{Bd} and \textit{Sp}, do not demonstrate any inhibitory activity (Figure~\ref{fig:ChInhib_NcraChytrid}B and C). Additional Chytridiomycota species (\textit{Operculomyces laminatus} JEL 223, \textit{Rhizoclosmatium hyalinus} JEL 800, and \textit{Obelidium mucronatum} JEL 802) do not inhibit hyphal growth of \textit{N. crassa} (Figure~\ref{fig:ChInhib_NcraChytrid}D-F), supporting the conclusion that this property is unique to \textit{Hp}. \\
\indent This phenomenon is not the result of any thigmotropism-related response or object avoidance type behavior in \textit{N. crassa}, as hyphae of \textit{N. crassa} are not inhibited by an agar block lacking \textit{Hp} sporangia (Figure~\ref{fig:ChInhib_NcAvoidance}).\\
\indent To determine if the sensitivity to \textit{Hp} sporangia was unique to \textit{N. crassa}, I screened a panel of fungi from among the Ascomycota, Basidiomycota, and Zygomycota. Figure~\ref{fig:ChInhib_HpOtherFungi}A recapitulates the previous observations about \textit{N. crassa}, while Figure~\ref{fig:ChInhib_HpOtherFungi}B and C demonstrate, respectively, that vegetative hyphal growth of \textit{N. discreta} and \textit{N. tetrasperma} can also be suppressed by the presence of \textit{Hp} sporangia.\\
\indent Within the Ascomycota, but outside of the genus Neurospora, \textit{T. reesei} (Sordariomycetes; Hypocreales; Hypocreaceae), \textit{A. nidulans} (Eurotiomycetes; Eurotiales; Trichocomaceae), and \textit{A. gossypii} (Saccharomycetes; Saccharomycetales; Saccharomycetaceae) are all completely sensitive to \textit{Hp} (Figure~\ref{fig:ChInhib_HpOtherFungi}D-F, respectively). \\
\indent Within the Basidiomycota, \textit{C. cinerea} (Agaricomycetes; Agaricales; Psathyrellaceae) is completely sensitive to \textit{Hp} sporangia (Figure~\ref{fig:ChInhib_HpOtherFungi}G). Growth of members of the order Mucorales had mixed sensitivities. \textit{P. blakesleeanus} (Mucorales; Phycomycetaceae) was completely sensitive to \textit{Hp} sporangia, yet \textit{R. oryzae} (Mucorales; Mucoraceae) appeared to be insensitive (Figure~\ref{fig:ChInhib_HpOtherFungi}H \& I, respectively). With a limited, but phylogenetically diverse panel of ten fungi, the broadly observed pattern of sensitivity suggests that the responsible compound has a generalized mechanism of action.\\
\indent Additionally, the growth in \textit{Hp} conditioned media filtrate of two strains of yeast (\textit{S. cerevisiae} MAU99 and AH109) and \textit{E. coli} DH5$\alpha$ suggested that this filtrate inhibited liquid growth of these organisms, and also provides evidence that the antimicrobial properties of \textit{Hp} extend to bacteria as well (Figure~\ref{fig:ChInhib_HpYeastEcoli}).\\

%---- Physiology ----%
\subsection*{The physiology of \textit{Hp} suggests that it is more tolerant of environmental stresses than \textit{Bd}}
To assess whether or not the observed growth suppression effect was the result of acidification of the local environment, \textit{Hp} was cultured on plates augmented with bromophenol blue or phenol red. After 96h growth, neither plates changed color, indicating that \textit{Hp} sporangia do not decrease the pH of mPmTG media below 6.8. However, phenol red plates which were additionally inoculated with \textit{N. crassa} did change from red to yellow in the areas around \textit{N. crassa} hyphae, suggesting that \textit{N. crassa} acidifies mPmTG media to below pH 6.8. This effect is not observed in bromophenol blue plates, suggesting that the final pH lies between 6.8 and 4.6.\\
\indent To assess the environmental tolerances of \textit{Hp} compared to \textit{Bd}, these two species were individually cultured in liquid media (mPmTG and 1\% Tryptone, respectively) prepared at different pHs, ranging from 4.0 to 9.0. The average OD$_{450}$, a measure of cell density, of \textit{Hp} at low pH (4.0 and 5.0) was 0.037 and 0.068 respectively, which was higher than that for \textit{Bd} (0.001 and 0.004 respectively). This suggests that \textit{Hp} is more tolerant of acidic environments than \textit{Bd} (Figure~\ref{fig:ChInhib_HpBdpHTemp}A). Similarly, \textit{Hp} and \textit{Bd} were grown at temperatures ranging from 18$^{\circ}$C to 30$^{\circ}$C, and the OD$_{450}$ assessed during a 21-day period (Figure~\ref{fig:ChInhib_HpBdpHTemp}B). Taken together, these results suggest that \textit{Hp} has a higher tolerance for environmental changes than \textit{Bd}. This finding is consistent with the fact that this isolate was obtained from an acidic lake \cite{Longcore2011}.\\

%---- Liquid Screen ----%
\subsection*{Liquid filtrate screening suggests a non-protein compound is responsible for bioactivity}
I was interested in obtaining a sample of bioactive liquid, on which I could perform chemical profiling to isolate a responsible compound. To test for bioactivity, \textit{Hp} cultured in liquid mPmTG media for a period of 72 hrs and subsequently filter-sterilized, producing "conditioned media" to which \textit{N. crassa} conidia was reintroduced. \\
\indent This conditioned media, obtained from \textit{Hp} and absent of all sporangia, retains the same inhibitory properties against \textit{N. crassa} observed with \textit{Hp} sporangia on solid agar plates (Figure~\ref{fig:ChInhib_FilterTempAssay}A). Similarly, \textit{N. crassa} was grown in mPmTG media, and after filter-sterilization, conditioned media was demonstrated to still provide sufficient sugar content to support growth of reintroduced \textit{N. crassa} conidia (Figure~\ref{fig:ChInhib_FilterTempAssay}B), supporting the idea of the presence of an active inhibitory compound rather than an effect of nutrient limitation. Additionally, conditioned media derived from "\textit{Hp} alone" and "\textit{Hp}+\textit{Nc}" preparations (Figure~\ref{fig:ChInhib_FilterTempAssay}A \& C) were indistinguishable from one another in their effects on \textit{N. crassa} growth, providing evidence to suggest that the compound is constitutively produced by \textit{Hp} and not as a response to a foreign competitor. \\
\indent The stability of the conditioned media was tested at seven different temperatures, ranging from -80$^{\circ}$C to 90$^{\circ}$C. The conditioned media retained bioactivity between -20$^{\circ}$C and 60$^{\circ}$C (Figure~\ref{fig:ChInhib_FilterTempAssay}D-E), but was ineffective after treatment of -80$^{\circ}$C and 90$^{\circ}$C (Figure~\ref{fig:ChInhib_FilterTempAssay}F-G). \textit{N. crassa} has no problem growing in fresh mPmTG media (Figure~\ref{fig:ChInhib_FilterTempAssay}H). Additionally, the conditioned media was ineffective if left sterile at room temperature for a period of 96h in the light, and subsequently inoculated with \textit{N. crassa} conidia.\\
\indent After treatment of 100 mg/ml of Proteinase K, the conditioned media was no less effective at inhibition of \textit{N. crassa} hyphal growth than untreated media. Taken together, these results suggest that a protein is not responsible for the observed bioactivity.\\
\indent A dilution series was prepared using VM media suggested that \textit{Hp} filtrate loses efficacy at 75\% of growth media.\\

%---- In silico analyses ----%
\subsection*{\textit{in silico} analyses}
Secondary metabolite clusters were predicted from the revised Illumina genome assembly using antiSMASH \cite{Blin2013} and SMURF \cite{Khaldi2010}. Using antiSMASH, several proteins in the \textit{Hp} genome were predicted to be likely homologs of terpene synthases. These genes were unique to \textit{Hp} when compared with other Chytridiomycota and Blastocladiomycota species. This analysis predicted three putative terpene synthase proteins. A FASTA search against the Illumina genome assembly recovered three additional proteins which did not appear in the second antiSMASH prediction. A list of the putative terpene synthase proteins from the Illumina assembly, along with their top BlastP results against the SwissProt database, is provided in Table~\ref{tab:ChInhib_Terpene}. The screen using antiSMASH also predicted 26 putative NRPS genes. \\
\indent Using SMURF, there were no putative secondary metabolite clusters predicted for \textit{Hp}. For comparison, 2 clusters were predicted in \textit{B. dendrobatidis}, 9 in \textit{S. punctatus}, and 2 in \textit{A. macrogynus}.\\
\indent To corroborate the experimental results which suggested the active molecule is not a protein, I performed \textit{in silico} secretome predictions using the available chytrid proteomes with a workflow described previously for assessing secretome prediction programs \cite{Min2010}. Table~\ref{tab:ChInhib_ChySec} lists the number of proteins in the examined chytrid genomes which appear to be secreted. In \textit{Hp}, 58 proteins were predicted to be secretome proteins. Using the secretome protein predictions from all surveyed chytrid genomes, OrthoMCL produced 350 ortholog clusters. However, none of these were unique to \textit{Hp}. Furthermore, none of the candidates were predicted to synthesize known antimicrobial or toxic proteins, which is supported by the temperature and proteinase treatments and further suggests that the compound is not enzymatic in nature.\\
\subsubsection*{Kinase mutant screen}
Finally, in order to address potential mechanisms of action for this behavior, an undergraduate student Na Jeong randomly assayed 2 \textit{N. crassa} phosphotase and 17 kinase mutants from \cite{Park2011}. All were found to be inhibited by growing sporangia of \textit{Hp} (Table~\ref{tab:ChInhib_Kinase}).\\

%%%%%%%%%%%%%%%%
%% Discussion %%
%%%%%%%%%%%%%%%%
\section{Discussion}
This study was an attempt to characterize a startling and unexpected observation: actively growing \textit{Hp} sporangia inhibit the vegetative, hyphal growth of \textit{N. crassa}. The findings demonstrate that i) \textit{Hp} is unique among surveyed chytrids in its ability to inhibit filamentous growth of \textit{N. crassa}, ii) the interaction is not limited to \textit{N. crassa} but is broadly active against a number of Ascomycete, Basidiomycete, and Zygomycete species, and iii) in silico secondary metabolite analysis suggests a terpene synthase-related enzyme as a possible candidate based on observed unique expansion of genes.\\
\indent \textit{Hp}, an aquatic, non-parasitic member of the Chytridiomycota isolated from a lake in the northeastern United States, and \textit{N. crassa}, a filamentous, multicellular, non-parasitic fungus primarily observed on decaying plant matter, are unlikely to come in specific contact with one another in the environment, so this response was presumed to be a non-specific interaction. The results from screening a phylogentically broad set of fungal species support this hypothesis, and it appears that compound(s) produced by \textit{Hp} are active against a multitude of fungal species.\\
\indent Compared with other chytrid isolates which were assayed, \textit{Hp} stands alone in its antifungal properties. Although the described basal lineages represent only 2\% of all described fungi, and the estimated diversity of these lineages is presumed high, this is the first time an interaction of this type has been described in any of these lineages. While it is possible that this behavior is not unique to this specific \textit{Hp} isolate, a more exhaustive assessment of chytrid isolates is necessary to elaborate on this hypothesis.\\
\indent The observation that conditioned media derived from "\textit{Hp} alone" and "\textit{Hp}+\textit{Nc}" preparations were indistinguishable from one another in their effects on \textit{N. crassa} growth suggests that the compound is being constitutively produced from \textit{Hp} and is not a response to a fungal competitor. Analysis of the transcriptome would be beneficial and would in theory easily identify any constitutively active genes in the sporangia life stage, from which this compound is expected to originate.\\
\indent The liquid compound profiling suggests that the compound is not protein-based, given that filter-sterilized conditioned media was a) effective after 65$^{\circ}$C treatment, a temperature at which many proteins denature, and b) effective after treatment of Proteinase K. For additional support for this hypthesis, I performed both \textit{in silico} secretome and secondary metabolite predictions. While the former did not predict any potential secreted proteins with known antimicrobial or antifungal properties, the latter methods predicted several terpene synthase related enzymes which appear to be uniquely present in \textit{Hp} compared to the other basal fungi proteomes searched.\\
\indent Taken together, these data tentatively suggest that the compound could be a constitutively-produced compound with broadly specific activity. While its mechanism of action is unknown, as are its chemical structure and related biosynthetic pathway(s), it is unlikely to be a protein or protein complex and is more likely to be a secondary metabolite. Future work will necessarily focus on chemical profiling of bioactive spent media filtrate to generate a working hypothesis for the chemical nature of the product. A starting point for this research is provided in the form of in silico genomic and transcriptomic research. Finally, it is worth noting that the relative ease with which this observation was made speaks to the necessity for further research into these basal lineages, of which intimate genomic and biochemical knowledge is lacking.\\
